On the Functionality of the Multilayer Perceptron

### 2020-12-14 Thesis/Drafts/Buell-PHYS799-ThesisDraft0.tex
Added v3, Finished abstract & intro

The task of recognizing the source of a sound wave is trivial and almost effortless to a human being.

@@

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place the data or a subset of the data into distinct categories called \textit{classes} based on it's properties. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to detect fraudulent transactions based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do unaided by some form of automation.

@@

@@

\paragraph*{}The design of the neural function that we have constructed combines two architecture types by using two distinct entry arrays which are processed, and a single output is formed. Thus type of neural network is referred to as a \textit{multimodal} nueral network \cite{Li,Ngiam}.

\paragraph*{}\textcolor{red}{This is not my best into to date}

@@

\item[•]Dakota Buell, Tan Dao, John Parker, Nathan Richard, Morgan Saidel for consultation on physics, mathematics, and programming topics.

>>>>

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being, but far more difficult for a computer to do. To execute this task, we have designed and constructed a multimodal neural network, that utilizes two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time discrete Fourier transformations (STDFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate the of the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates an average superior classification performance when compared to either of it's constituent unimodal architectures.

@@

\paragraph*{}The ability to classify time-series data into unique categories has become topic of much attention in the last few decades \cite{Khan,Mierswa}. From climate measurements, to financial reports, to audio-visual media, or to medicine, time-domain signals surround the modern world at every corner \cite{Virtanen}. In each context, the general idea is to place data or a subset of data into distinct categories called \textit{classes} based on it's properties. For example, doctors may wish to classify a patient as having a heart condition based on properties of their heartbeat or banks may wish to classify a purchase as fraudulent based on a history of spending habits. Often, the volume  of information or complexity of the task makes it unreasonable for a human to do unaided by some form of automation.

@@

\paragraph*{}As the name implies, a neural network is a computation a model or graph that is loosely structured off of the human \cite{Geron,Loy}. Where a biological brain uses physical connects to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a mathematical model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years, later, one of the earliest computation neural networks called the \textit{perceptron} was developed by Frank Rosenblatt in 1953, which modeled connections of neurons through principles of linear algebra \cite{Geron}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

\paragraph*{}Modern developments in high-performance computing and

@@

\paragraph*{}The design of our neural function combines two architecture types by using two distinct entry arrays which are processed, and a single output is formed. Thus type of neural network is referred to as a \textit{multimodal} nueral network \cite{Li,Ngiam}. Our multimodal network contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is it's own computuational graph that is designed to process it's respective inputs types and

@@

\item[•]Dakota Buell, Tan Dao, John Parker, Nathan Richard, and Morgan Saidel for consultation on physics, mathematics, and programming topics.

### 2020-12-14 Thesis/Drafts/Buell-PHYS799-ThesisDraft3.tex
Added v3, Finished abstract & intro

\begin{titlepage}
	\begin{center}
   		\vspace*{1cm}
       	\begin{LARGE}
       	\textbf{Mapping Chaotically Generated Synthesizers to Musical Instruments with Machine Learning and Harmonic Analysis} \\
       	\end{LARGE}
       	\vspace{1cm}
       	\begin{Large}
       	Multimodal Classification of Digital Signal
       	\end{Large}             
       	\vspace{1.5cm}

@@

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being, but far more difficult for a computer to do. To execute this task, we have designed and constructed a multimodal neural network, that utilizes two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time discrete Fourier transformations (STDFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate the of the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates an average superior classification performance when compared to either of it's constituent unimodal architectures.

@@

\paragraph*{}It has long been a goal of computer scientists and engineers alike to develop a program, algorithm, or machine that could perform a certain task to the level of human-being \cite{Bishop,Geron,Loy}. While humans are intricate organisms capable of recognition or classification in many contexts, we are not notorious for the ability to process millions of raw numerical values in the course of a few seconds. Similarly, modern computers are also complex systems that can filter through exabytes of information on a daily basis, but are not generally known for being able to distinguish cats from dogs \cite{Geron}. The compromise between is an algorithm called a \textit{neural network} that combines the computation speed of a computer with the decision-making power of a human \cite{Goodfellow,Levine,Loy}.

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instrument that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Some papers details features and performances of attempting to classify speech against music or organize audio-visual media based on content.

\paragraph*{}As the name implies, a neural network is a computation a model or graph that is loosely structured off of the human \cite{Geron,Loy}. Where a biological brain uses physical connects to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a mathematical model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years, later, one of the earliest computation neural networks called the \textit{perceptron} was developed by Frank Rosenblatt in 1953, which modeled connections of neurons through principles of linear algebra \cite{Geron}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

\paragraph*{}Modern developments in high-performance computing and

\paragraph*{}It is well know that the choice of inputs, called features is critically important to classification success \cite{Khan,Serizel}. In many cases, we can develop and explore a set of features that allows for the particular classification task to be executed considerably well. In our case, the choice of features is motivated by the physics of each musical instrument class. By exploring the nature of the vibration part of each instrument according to the Hornbostel-Sachs classification system, we can describe some of the mechanical properties and see how they differ between instruments of different types. These differences are key to providing a set of features which allow a neural network to differentiate between classes of instruments \cite{Liu,Zhang}.

\paragraph*{}The design of our neural function combines two architecture types by using two distinct entry arrays which are processed, and a single output is formed. Thus type of neural network is referred to as a \textit{multimodal} nueral network \cite{Li,Ngiam}. Our multimodal network contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is it's own computuational graph that is designed to process it's respective inputs types and

@@

\paragraph*{}A Maximum Pooling layers returns the maximum neuron activation in a two-dimensional group of neurons. In the case of 2D Max Pooling, we choose a kernel size to be $m \times n$, similar to the convolution layer, and extract the maximum value in each window, while stepping along according to a chosen stride size \cite{Loy,Goodfellow}. Consider an input like in Fig. (\ref{fig-2DConvExample}a). Using the $2 \times 2$ kernel on the $3 \times 4$ input, each box would then contain the maximum value of the input activations. We detail this in Fig. (\ref{fig-2DMaxPool}):
\begin{figure}[H]
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(d)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}
\caption{The result of 2D maximum-pooling an input array. This image was adapted from Loy, pg. 126 \cite{Loy}}
\label{fig-2DMaxPool}
\end{figure}

@@

\paragraph*{}We can carry this idea over by producing a \textit{multimodal} neural network that accepts two different forms of input for each sample. One input is an $N' \times k$ spectrogram matrix, explored in sec. (\ref{subsec-Spectrogram}) and the other input is a $1 \times p$ feature vector explored in sec. (\ref{subsec-TimeFeatures}) and sec. (\ref{subsec-FreqFeatures}). The spectrogram acts to provide an energy distribution of a signal as a function of frequency and time, while the feature vector provides qualities in a more list-like fashion. This type of supervised learning is referred to as \textit{multi-view} learning, and we detail a conceptual diagram of the architecture in Fig. (\ref{fig-NetworkArchitecture}) \cite{Li,Ngiam}.

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.75]{../Figures/NeuralNetworkArchitecture}
\end{center}
\caption{The implemented miltimodal architecture of the audio file classification neural network. The Left branch process an image-like input, the right branch processes a vector-like input. The activations are then merged, and then a single output vector is produced}
\label{fig-NetworkArchitecture}
\end{figure}

@@

\subsubsection{The Spectrogram Branch}

\paragraph*{}The spectrogram branch is pictured on the left side of Fig. (\ref{fig-NetworkArchitecture}), and is an implementation of a \textit{convolutional neural networl} (CNN). A spectrogram is a representation of a sound wave, or signal by representing energy distributions as a function of \textit{time} and \textit{frequency} \cite{White,Olson,Khan}. The input layer of the spectrogram accepts a 4-Dimensional array. The axes, in order of indexing, represents (i) the size of the \textit{mini-batch} of samples, (ii) the pixel height of each sample, (iii) the pixel wifth of each sample, and (iv) the number of channels in the image. For this model, we have selected to use $64$ samples per batch, $558$ pixels in height, $256$ pixels in width, and $1$ gray-scale channel. We denote the 4D shape of the design matrix for this branch as:
\begin{equation}
\label{eqn-shapeX1}
X_1 \in \mathbb{R}^{(64 \times 558 \times 256 \times 1)}
\end{equation}
Any other shape will be rejected, and an error is raised.

@@

\paragraph*{}The matrix $S$, is the spectrogram representation of the initial waveform, and has shape $N \times k$ of all real floating-point numbers. We can index matrix $S$ similarly to that of matrix $A$ in Eq. (\ref{eqn-IndexingA}):
\begin{equation}
\label{eqn-IndexingS}
S_{i,j} = S_i[j] = S[i,j]
\end{equation}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.3]{../Figures/LossHistory}
\end{center}
\caption{The loss function score decreases with each training step, indicating that optimization is performing correctly}
\label{fig-LossScore}
\end{figure}
similarly, we can visualize the precision score and recall score at each training step.
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.3]{../Figures/MetricHistory}
\end{center}
\caption{The precision score and recall score generally increase with each training step, indicating that the model gis getting both more sensistive and more specific with each step}
\label{fig-PrecisionRecallScores}
\end{figure}

@@

\subsection{Predictions on Chaotic Synthesizers}

\textcolor{red}{Help}

>>>>

\begin{titlepage}
	\begin{center}
   		\vspace*{1cm}
       	\begin{LARGE}
       	\textbf{Audio Classification of Musical Instruments with a Multimodal Neural Network} \\
       	\end{LARGE}
       	\vspace{6cm}

@@

The task of recognizing the source of a sound wave in a digital audio file is trivial and almost effortless to a human being, but far more difficult for a computer to do. To execute this task, we have designed and constructed a multimodal neural network, that utilizes two different input branches, one being a convolutional neural network (CNN), and the other, a multilayer perceptron (MLP). The CNN branch accepts input in the form of a two-dimensional spectrogram matrix, which encodes the energy distribution of an audio signal over frequency and time, and the MLP branch accepts input as a one-dimensional array of features derived from the physical nature of the sound wave. The spectrogram is produced by using a concatenation of multiple short-time Fourier transformations (STFT) of overlapping windowed time-space analysis frames. The feature vector implements techniques from digital signal processing to combine more explicit properties of the time-space and frequency-space information. To classify the audio files based on predicted sources, we aggregate the of the activations from the final layers of each branch, and produce a single prediction label. We find that this architecture allows for reasonable classification performance, and demonstrates superior classification performance on average, when compared to either of it's constituent unimodal architectures. Training data for this experiment is from studio recordings of the Philharmonia Symphony Orchestra and University of Iowa's Electronic Music Studios.

@@

\paragraph*{}It has long been a goal of computer scientists and engineers alike to develop a program, algorithm, or machine that could perform a certain task to the level of human-being \cite{Bishop,Geron,Loy}. While humans are intricate organisms capable of recognition or classification in many contexts, we are not notorious for the ability to process millions of raw numerical values in the course of a few seconds. Similarly, modern computers are also complex systems that can filter through exabytes of information on a daily basis, but are not generally known for being able to distinguish cats from dogs \cite{Geron}. The compromise between humans and machines an algorithm called a \textit{neural network} that combines the computation speed of a computer with the decision-making power modeled from human brain \cite{Goodfellow,Levine,Loy}.

\paragraph*{}In this work, we have developed an automated neural network program that can classify digital sound waves as one of $37$ musical instruments that it most closely resembles. The general use of machine learning algorithms for audio information classification or identification is well explored, and a topic of much modern research \cite{Khan,Liu,Mierswa,Serizel,Zhang}. From previous work, it has been found that various types of audio classification are possible, and are still being expanded upon. Many papers details predictors for classifying speech against music or features for organizing audio-visual media based on content. In each case, features are often tailored to meet the specific task at hand \cite{Virtan,Liu}.

\paragraph*{}It has been experimentally determined that the choice of features is critically important to classification success \cite{Khan,Serizel}. In our case, the nature of these features is motivated by the physics and mechanical properties of each musical instrument class. By exploring the physics of the vibrating parts of each instrument according to the Hornbostel-Sachs classification system, we gain a more through understanding of each instrument group. This allows us to develop properties within each class that are consistent with a low variability, and what properties differ between classes and show a higher variability. These differences are key to providing a set of features which allow a neural network to differentiate between classes of musical instruments \cite{Liu,Zhang}.

\paragraph*{}As the name implies, a neural network is a computation a model or graph that is loosely structured off of the biological brain \cite{Geron,Loy}. Where a biological brain uses physical connects through axons to exchanges chemicals to communicate information, a neural network uses a chain of mathematical operations to communicate information \cite{Goodfellow}. The earliest work on attempting to model cognitive processes goes back to 1943, where Warren McCulloch and Walter Pitts began to develop a model of the human brain using principles of mathematics \cite{McCulloch,Loy}. Nearly ten years later, a simple neural networks called the \textit{perceptron} was developed by Frank Rosenblatt, which modeled connections of neurons through principles of linear algebra \cite{Geron}. Another decade after that, inspiration from the visual cortex gave rise to the convolution operation to be applied to digital image processing.

\paragraph*{}The design of our neural network function combines two architecture types by using two distinct entry arrays which are processed, and a single output is formed. This type of neural network is referred to as a \textit{multimodal} neural network \cite{Li,Ngiam}. Our model contains a convolutional neural network (CNN) that processes an image-like representation of an audio waveform, and a multilayer perceptron (MLP) that processes a list-like representation of an audio wave form. Each arm of the network is it's own computational graph that is designed to process it's respective inputs types. The activations at the end of each layer are concatenated into a new layer, which is further processed and then used to generate a single prediction. This prediction aggregated from two features on average shows superior classification when compared to the performance either of either of it's constituent unimodal architectures alone.

@@

\paragraph*{}A Maximum Pooling layers returns the maximum neuron activation in a two-dimensional group of neurons. In the case of 2D Max Pooling, we choose a kernel size to be $m \times n$, similar to the convolution layer, and extract the maximum value in each window, while stepping along according to a chosen stride size \cite{Loy,Goodfellow}. Consider an input like in Fig. (\ref{fig-2DConvExample}a). Using the $2 \times 2$ kernel on the $3 \times 4$ input, each box would then contain the maximum value of the input activations. We detail this in Fig. (\ref{fig-2DMaxPool}):
\begin{figure}[H]
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(a)}
	\caption{Input Activations, $x^{(l-1)}$}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
	\includegraphics[scale=0.6]{../figures/2DConvExample(d)}
	\caption{Output Activations, $x^{(l)}$}
	\end{subfigure}
\caption{The result of 2D maximum-pooling an input array. This image was adapted from Loy, pg. 126 \cite{Loy}}
\label{fig-2DMaxPool}
\end{figure}

@@

\paragraph*{}We can carry this idea over by producing a \textit{multimodal} neural network that accepts two different forms of input for each sample. One input is an $N' \times k$ spectrogram matrix, explored in sec. (\ref{subsec-Spectrogram}) and the other input is a $1 \times p$ feature vector explored in sec. (\ref{subsec-TimeFeatures}) and sec. (\ref{subsec-FreqFeatures}). The spectrogram acts to provide an energy distribution of a signal as a function of frequency and time, while the feature vector provides qualities in a more list-like fashion. This type of supervised learning is referred to as \textit{multi-representation learning}, and we detail a conceptual diagram of the architecture in Fig. (\ref{fig-NetworkArchitecture}) \cite{Li,Ngiam}.

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.75]{../Figures/NeuralNetworkArchitecture}
\end{center}
\caption{The implemented multimodal architecture of the audio file classification neural network. The left branch process an image-like input, and the right branch processes a vector-like input. The activations are then merged, and then a single output vector is produced}
\label{fig-NetworkArchitecture}
\end{figure}

@@

\subsubsection{The Convolution Branch}

\paragraph*{}The convolution branch is pictured on the left side of Fig. (\ref{fig-NetworkArchitecture}), and is an implementation of a \textit{convolutional neural network} (CNN). It processes a spectrogram, which is a representation of a waveform, or signal as an energy distributions over \textit{time} and \textit{frequency} \cite{White,Olson,Khan}. The input layer of this branch accepts a 4-Dimensional array. The axes, in order of indexing, represents (i) the size of the \textit{mini-batch} of samples, (ii) the pixel height of each sample, (iii) the pixel wifth of each sample, and (iv) the number of channels in the image. For this model, we have selected to use $64$ samples per batch, $558$ pixels in height, $256$ pixels in width, and $1$ gray-scale channel. We denote the 4D shape of the design matrix for this branch as:
\begin{equation}
\label{eqn-shapeX1}
X_1 \in \mathbb{R}^{(64 \times 558 \times 256 \times 1)}
\end{equation}
Any other shape will be rejected, and an error is raised.

@@

\paragraph*{}The matrix $S$, is the spectrogram representation of the initial waveform, and has shape $N \times k$ of all real floating-point numbers. We can index matrix $S$ similarly to that of matrix $A$ in Eq. (\ref{eqn-IndexingA}):
\begin{equation}
\label{eqn-IndexingS}
S_{i,j} = S_i[j] = S[i,j]
\end{equation}
Note that the spectrogram provided to the neural network is representing the power spectrum of the audio file as defined in Eq. (\ref{eqn-Spectrogram}). All spectrogram figures in this report are from taking the element-wise \textit{natural log} of the power spectrum. This has been done solely for visualization purposes.

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.3]{../Figures/LossHistory}
\end{center}
\caption{The loss function score decreases with each training step, indicating that optimization is performing correctly}
\label{fig-LossScore}
\end{figure}
similarly, we can visualize the precision score and recall score at each training step.
\begin{figure}[H]
\begin{center}
\includegraphics[scale=0.3]{../Figures/MetricHistory}
\end{center}
\caption{The precision score and recall score generally increase with each training step, indicating that the model is getting both more sensitive and more specific as training progresses}
\label{fig-PrecisionRecallScores}
\end{figure}

@@
